Toward single particle reconstruction without particle picking: Breaking the detection limit

### 2018-09-03 pnas/Towards.tex
Update appendices B-D

\begin{proposition} \label{prop:gamma_sigma}
	%
	%Let $\sigma > 0$ be fixed. 
	The observed autocorrelations $a_y^1,a_y^2$ and  $a_y^3$ determine the ratio $\gamma$ and noise level $\sigma$ uniquely for a generic signal $x$. If $\gamma\geq\frac{L}{4(L-1)}\approx \frac{1}{4}$, then this holds for any signal $x$ with nonzero mean. 
	\begin{proof}
		See Appendix~\ref{sec:proof_prop_gamma_sigma}.
	\end{proof}
\end{proposition}

From Propositions~\ref{prop:uniqueness} and~\ref{prop:gamma_sigma} we can directly deduce the following:
\begin{corollary}
	%Let $K=1$ and $\sigma > 0$ be fixed.
	 The signal $x$, the ratio $\gamma$, and the noise level $\sigma$ are identifiable from the first three autocorrelation functions of $y$ if:
	\begin{itemize}
		\item Either the signal $x$ is generic; or
		\item   $x$ has nonzero mean  and $\gamma\geq\frac{L}{4(L-1)}$.
	\end{itemize}
\end{corollary}
%
%\TODO{Similar results hold for signals in higher dimensions.---need to keep this sentence in red until the very end to make sure it's correct.}

@@

%Throughout the proof, we consider the case of one signal $K=1$. The extension to $K>1$ is straightforward by averaging the contributions of all signal with  appropriate weights; see~\cite{boumal2017heterogeneous}. 
%We will let the number of instances of the signal $M$ grow with $N$, and write $M=M_N$ to emphasize this.
We consider the asymptotic regime $N,M\to\infty$ and assume that $M=\Omega(N)$, so that
% We assume $M_N$ grows proportionally with $N$, and recall that
% define: %\TODO{---Nico: the paper now no longer references $M_N$ and defines $\gamma$ 'as is' for given $M$ and $N$; may want to adapt things here as a result. It wouldn't make sense to make $M_N$ appear in the paper I think, as it just makes things more complicated. In a sense, much of this beginning of appendix is just repeating things we already said cleanly in the paper: let's just follow up from that.}
%
\begin{equation*}
\gamma := \frac{ML}{N}>0.
\end{equation*}

@@

We proceed with the (second-order) autocorrelation for fixed $\ell\in[0,\ldots,L-1]$. We can compute:
%
\begin{align*}
%
a_y^2[\ell] & = \frac{1}{N}\sum_{i=0}^{N-1-\ell}y[i]y[i+\ell]
\nonumber \\
& = \underbrace{\frac{1}{N}\sum_{j=1}^{M}\sum_{i=0}^{L-\ell-1}x[i]x[i+\ell]}_{\text{signal term}} + \underbrace{\frac{1}{N}\sum_{i=0}^{N-1-\ell}\varepsilon[i]\varepsilon[i+\ell]}_{\text{noise term}}
\\ & + \underbrace{\frac{2}{N} \sum_{j=1}^{M} \sum_{i=0}^{L-1} x[i] \varepsilon[s_j + i + \ell]}_{\text{cross-term}}. 
%
\end{align*}

The cross-terms are linear in the noise, and are easily shown to vanish almost surely in the limit $N\to\infty$, by the strong law of large numbers. As for the signal term, we break it into $M$ different sums, each containing one copy of the signal. This gives:
%
\begin{align} \label{eq:2nd_moment_signal_term}
%
\frac{1}{N}\sum_{j=1}^{M}\sum_{i=0}^{L-\ell-1}x[i]x[i+\ell] & = \frac{ML}{N}\frac{1}{L}\sum_{i=0}^{L-\ell-1}x[i]x[i+\ell]\nonumber\\&\xrightarrow{N\to\infty}\gamma a_x^2[\ell].
%
\end{align}
%

We next analyze the pure noise term. When $\ell\neq 0$, we can break the noise term into a sum of independent terms:
%
\begin{equation}
%
\frac{1}{N}\sum_{i=0}^{N-1-\ell} \varepsilon[i]\varepsilon[i+\ell] = \frac{1}{\ell}\sum_{i=0}^{\ell-1}\frac{1}{N/\ell}\sum_{j=0}^{N/\ell -1} \varepsilon[j\ell + i] \varepsilon[(j+1)\ell + i].
%
\end{equation}
%
Each sum $\frac{1}{N/\ell}\sum_{j=0}^{N/\ell -1} \varepsilon[j\ell + i] \varepsilon[(j+1)\ell + i]$ is an average of $N/\ell$ independent terms with expectation zero, and thus converges to zero almost surely as $N\to\infty$. If $\ell=0$, then we have:
%
\begin{equation*}
%
\frac{1}{N}\sum_{i=0}^{N-1} \varepsilon^2[i] \xrightarrow{a.s.} \sigma^2.
%
\end{equation*}

We now analyze the third-order autocorrelation. Let us fix $\ell_1\geq\ell_2\ge0$. We have:
%
\begin{align*}
%
&a_y^3[\ell_1,\ell_2] 
= \frac{1}{N}\sum_{i=0}^{N-1-\ell_1} y[i]y[i+\ell_1]y[i+\ell_2]
\nonumber \\
%
=& \underbrace{ \frac{ML}{N}\frac{1}{M}\sum_{j=1}^M 
	\frac{1}{L}\sum_{i=0}^{L-1-\ell_1}x[i]x[i+\ell_1]x[i+\ell_2]   }_{(1)}
\\& + \underbrace{\frac{1}{N}\sum_{i=0}^{N-1-\ell_1} \ep[i]\ep[i+\ell_1]\ep[i+\ell_2]}_{(2)}
\nonumber \\
&+ \underbrace{\frac{1}{N}\sum_{j=1}^{M} 
	\sum_{i=0}^{L-1} x[i]\ep[s_j + i+\ell_1]\ep[s_j+ i+\ell_2]}_{(3)}
\\& + \underbrace{\frac{1}{N}\sum_{j=1}^{M} 
	\sum_{i=0}^{L-1} \ep[s_j+i-\ell_1]x[i]\ep[s_j+ i+\ell_2-\ell_1]}_{(4)}
\nonumber \\
&+ \underbrace{\frac{1}{N}\sum_{j=1}^{M} 
	\sum_{i=0}^{L-1} \ep[s_j+i-\ell_2]\ep[s_j+i+\ell_1-\ell_2]x[i]}_{(5)}
\\ & + \underbrace{\frac{1}{N}\sum_{j=1}^{M} 
	\sum_{i=0}^{L-\ell_1+\ell_2-1} \ep[s_j+i]x[i+\ell_1-\ell_2]x[i]}_{(6)}
\nonumber \\
&+ \underbrace{\frac{1}{N}\sum_{j=1}^{M} 
	\sum_{i=0}^{L-\ell_2-1} x[i]\ep[s_j + i+\ell_1]x[s_j+ i+\ell_2]}_{(7)}
\\ & + \underbrace{\frac{1}{N}\sum_{j=1}^{M} 
	\sum_{i=0}^{L-\ell_1-1} x[i]x[i+\ell_1]\ep[s_j+ i+\ell_2]}_{(8)}.
%
\end{align*}
%
Terms (6), (7) and (8) are linear in $\ep$, and can easily be shown to converge to 0 almost surely by the law of large numbers, by similar arguments as used previously. Term (1) converges to $\gamma a_x^3[\ell_1,\ell_2]$ almost surely, for the same reasons as~\eqref{eq:2nd_moment_signal_term}. To deal with terms (2)--(5), we must distinguish between different values of $\ell_1$ and $\ell_2$.

@@

{\bf Case 2:} $0=\ell_2 < \ell_1$. Term (2) is an average of products of the form $\ep[i]^2\ep[i+\ell_1]$, which have mean zero; consequently, term (2) converges to 0 almost surely. The same argument as for Case 1 shows that (3) and (5) also converge to 0. For term (4), we write:
%
\begin{align*}
%
&\frac{1}{N}\sum_{j=1}^{M} 
\sum_{i=0}^{L-1} \ep[s_j+i-\ell_1]x[i]\ep[s_j+ i+\ell_2-\ell_1]
\nonumber \\
&= \frac{ML}{N}\frac{1}{L}\sum_{i=0}^{L-1}x[i] \frac{1}{M}\sum_{j=1}^{M} \ep[s_j+i-\ell_1]^2
\\& \xrightarrow{N\to\infty} \gamma \frac{1}{L} \sum_{i=0}^{L-1}x[i] \sigma^2 = \gamma a_x^1 \sigma^2. \nonumber
%
\end{align*}

@@

Now let us compute the probability of failure conditional on the event $\eta = 0$. In this case, failure occurs when $X \in S$. Since $X|(\eta=0) \sim N(\theta_0,\sigma^2)$, we can write $X | (\eta=0) = \sigma Z + \theta_0$, where $Z \sim N(0,1)$. On that condition,

@@

\section{Proof of Proposition~\ref{prop:gamma_sigma}} \label{sec:proof_prop_gamma_sigma}
\TODO{Need to check the details of the proof one more time}
We will prove that both $\sigma$ and $\gamma$ are identifiable from the observed first three moments of $y$. For convenience, we will work with $\beta = \gamma / L$ rather than $\gamma$ itself. We will construct two quadratic equations satisfied by $\beta$ from observed quantities, independent of $\sigma$. Then, we will show that these equations are independent, and hence that $\beta$ is uniquely defined.  Given $\beta$, we can estimate $\sigma$ using Proposition~\ref{prop:gamma}.

@@

Recall that $a_y^1 = \beta(\one^Tx)$ and  
and $a_y^2[0] = \beta\|x\|^2+\sigma^2$, where $\one\in\RL$ stands for vector of ones. Taking the product:
\begin{equation}\label{eq:E1}
\begin{split}
E_1 &:= a_y^1a_y^2[0] =  (\beta(\one^Tx))(\beta\|x\|^2+\sigma^2) \\
& = \sigma^2a_y^1 + \beta^2F_1,
\end{split}
\end{equation}
where $F_1 := a_x^3[0,0] + \sum_{j=1}^{L-1}(a_x^3[j,j] + a_x^3[0,j])$. 
The terms of $F_1$ can be also estimated from $a_y^3$, while taking the scaling and bias terms into account:
\begin{equation} \label{eq:E2}
E_2:= \beta F_1 + (2L+1)\sigma^2a_y^1.
\end{equation}
Therefore, from~\eqref{eq:E1} and~\eqref{eq:E2} we get
\begin{equation} \label{eq:E12}
E_2\beta -(2L+1)\sigma^2\beta a_y^1 = E_1-\sigma^2a_y^1.
\end{equation}
Let $a_y^2:=a_y^2[0] + 2\sum_{j = 1}^{L-1}a_y^2[j]$ and recall from Proposition~\ref{prop:gamma}:
\begin{equation} \label{eq:sigma2}
\sigma^2 = a_y^2 - (a^1_y)^2/\beta. 
\end{equation} 
Plugging into~\eqref{eq:E12} and rearranging we get 
\begin{equation} \label{eq:quad1}
\mathcal{A}\beta^2 + \mathcal{B}\beta + \mathcal{C} = 0,
\end{equation}
where 
\begin{align*}
\mathcal{A} &= E_2 - (2L+1)a_y^1a_y^2, \\ 
\mathcal{B} &= -E_1 + (2L+1)(a_y^1)^3 + a_y^1a_y^2  , \\
\mathcal{C} &= -(a_y^1)^3.
\end{align*}
Importantly, these coefficients are observable quantities.

We are now proceeding to derive the second quadratic equation. We notice that 
\begin{equation} \label{eq:E3}
E_3  = \frac{1}{L}(a_y^1)^3 = \frac{1}{L}\beta^3 (\one ^Tx)^3   = \frac{1}{L}\beta^3 F_2,
\end{equation}
where 
\begin{equation*}
F_2 =  a_x^3[0,0] + 3\sum_{j=1}^{L-1}a_x^3[j,j] + 3\sum_{j=1}^{L-1}a_x^3[0,j] + 6\sum_{1\leq i < j\leq L-1}a_x^3[i,j].
\end{equation*}
On the other hand, from $a_y^3$ we can directly estimate $F_2$ up to scale and bias
\begin{equation*} \label{eq:E4}
E_4 = \beta F_2 + (6L-3)\sigma^2a_y^1.
\end{equation*}
Taking the ratio:
\begin{equation*} 
\frac{E_4}{E_3} = \frac{L}{\beta^2} + \frac{(6L-3)L\sigma^2a_y^1}{E_3}, 
\end{equation*}
we conclude:
\begin{equation*}
\sigma^2 = \frac{E_4}{a_y^1L(6L-3)}  - \frac{E_3}{\beta^2a_y^1(6L-3)}.
\end{equation*}
Using~\eqref{eq:sigma2} and rearranging we get the second quadratic:
\begin{equation} \label{eq:quad2}
\mathcal{D}\beta^2 + \mathcal{E}\beta + \mathcal{F} = 0,
\end{equation}
where
\begin{align*}
\mathcal{D} &= a_y^2 - \frac{E_4}{a_y^1L(6L-3)}, \\ 
\mathcal{E} &= -(a_y^1)^2, \\
\mathcal{F} &= \frac{E_3}{a_y^1(6L-3)}.
\end{align*}

To complete the proof, we need to show that the two quadratic equations~\eqref{eq:quad1} and~\eqref{eq:quad2} are independent. To this end, it is enough to show that the ratio between the coefficients is not the same. 
From~\eqref{eq:quad1} and~\eqref{eq:E1}, we have 
\begin{equation*}
\begin{split}
\frac{\mathcal{B}}{\mathcal{C}} &= \frac{LE_1 - (2L+1)(a_y^1)^3 - La_y^1a_y^2}{(a_y^1)^3} \\&= \frac{La_y^2[0] - (2L+1)(a_y^1)^2 - La_y^2}{(a_y^1)^2}.
\end{split}
\end{equation*}
In addition, using~\eqref{eq:E3}
\begin{equation*}
\frac{\mathcal{E}}{\mathcal{F}} = \frac{(3-6L)(a_y^1)^3}{LE_3} = 3 - 6L . 
\end{equation*}

Now, suppose that the quadratics are dependent. Then, $\frac{\mathcal{B}}{\mathcal{C}} =\frac{\mathcal{E}}{\mathcal{F}} $, or, 	
\begin{equation*}
La_y^2[0] - (2L+1)(a_y^1)^2 - La_y^2 = (a_y^1)^2(3-6L).
\end{equation*}
Rearranging the equation and writing in terms of $x$ we get 
\begin{equation} \label{eq:cond}
4(L-1)\beta (a_x^1)^2  - \sum_{i=1}^{L-1} a_x^2[i] = 0.
\end{equation}	
For generic $x$,  this polynomial equation is not satisfied. Therefore,  the equations are independent. 
More than that, for any nonzero $x$, $(a_x^1)^2 >\sum_{i=1}^{L-1} a_x^2[i]$. Therefore, if $4(L-1)\beta \geq 1$, or,
\begin{equation*}
\beta \geq \frac{1}{4(L-1)},
\end{equation*}
the condition~\eqref{eq:cond} cannot be satisfied for any signal.

>>>>

\begin{proposition} \label{prop:gamma_sigma}
	%
	%Let $\sigma > 0$ be fixed. 
	The observed autocorrelations $a_y^1,a_y^2$ and  $a_y^3$ determine the ratio $\gamma$ and noise level $\sigma$ uniquely for a generic signal $x$. If $\gamma\geq \frac{1}{4}$, then this holds for any signal $x$ with nonzero mean. 
	\begin{proof}
		See Appendix~\ref{sec:proof_prop_gamma_sigma}.
	\end{proof}
\end{proposition}

From Propositions~\ref{prop:uniqueness} and~\ref{prop:gamma_sigma} we can directly deduce the following:
\begin{corollary}
	%Let $K=1$ and $\sigma > 0$ be fixed.
	 The signal $x$, the ratio $\gamma$, and the noise level $\sigma$ are identifiable from the first three autocorrelation functions of $y$ if:
	\begin{itemize}
		\item Either the signal $x$ is generic; or
		\item   $x$ has nonzero mean  and $\gamma\geq\frac{1}{4}$.
	\end{itemize}
\end{corollary}
%
%\TODO{Similar results hold for signals in higher dimensions.---need to keep this sentence in red until the very end to make sure it's correct.}

@@

%Throughout the proof, we consider the case of one signal $K=1$. The extension to $K>1$ is straightforward by averaging the contributions of all signal with  appropriate weights; see~\cite{boumal2017heterogeneous}. 
%We will let the number of instances of the signal $M$ grow with $N$, and write $M=M_N$ to emphasize this.
We consider the asymptotic regime $N,M\to\infty$ and assume that $M=\Omega(N)$, so that
% We assume $M_N$ grows proportionally with $N$, and recall that
% define: %\TODO{---Nico: the paper now no longer references $M_N$ and defines $\gamma$ 'as is' for given $M$ and $N$; may want to adapt things here as a result. It wouldn't make sense to make $M_N$ appear in the paper I think, as it just makes things more complicated. In a sense, much of this beginning of appendix is just repeating things we already said cleanly in the paper: let's just follow up from that.}
%
\begin{equation*}
\gamma := \frac{ML}{N}>0.
\end{equation*}
%
Any vectors with indices out of range are given value 0.

@@

We proceed with the (second-order) autocorrelation for fixed $\ell\in[0,\ldots,L-1]$. We can compute:
%
\begin{align*}
%
a_y^2[\ell] & = \frac{1}{N}\sum_{i=0}^{N-1-\ell}y[i]y[i+\ell]
\nonumber \\
& = \underbrace{\frac{1}{N}\sum_{j=1}^{M}\sum_{i=0}^{L-\ell-1}x[i]x[i+\ell]}_{\text{signal term}} + \underbrace{\frac{1}{N}\sum_{i=0}^{N-1-\ell}\varepsilon[i]\varepsilon[i+\ell]}_{\text{noise term}}
\\ & + \underbrace{\frac{1}{N} \sum_{j=1}^{M} \sum_{i=0}^{L-1} x[i] (\varepsilon[s_j + i + \ell] + \varepsilon[s_j + i - \ell])}_{\text{cross-term}}. 
%
\end{align*}

The cross-term is linear in the noise, and is easily shown to vanish almost surely in the limit $N\to\infty$, by the strong law of large numbers. We break the signal term into $M$ different sums, each containing one copy of the signal. This gives:
%
\begin{align} \label{eq:2nd_moment_signal_term}
%
\frac{1}{N}\sum_{j=1}^{M}\sum_{i=0}^{L-\ell-1}x[i]x[i+\ell] & = \frac{ML}{N}\frac{1}{L}\sum_{i=0}^{L-\ell-1}x[i]x[i+\ell]\nonumber\\&\xrightarrow{N\to\infty}\gamma a_x^2[\ell].
%
\end{align}
%

We next analyze the pure noise term. When $\ell\neq 0$, we can break the noise term into a sum of independent terms:
%
\begin{equation*}
%
\frac{1}{N}\sum_{i=0}^{N-1-\ell} \varepsilon[i]\varepsilon[i+\ell] = \frac{1}{\ell}\sum_{i=0}^{\ell-1}\frac{1}{N/\ell}\sum_{j=0}^{N/\ell -2} \varepsilon[j\ell + i] \varepsilon[(j+1)\ell + i].
%
\end{equation*}
%
Each sum $\frac{1}{N/\ell}\sum_{j=0}^{N/\ell -2} \varepsilon[j\ell + i] \varepsilon[(j+1)\ell + i]$ is an average of $N/\ell$ independent terms with expectation zero, hence converges to zero almost surely as $N\to\infty$. If $\ell=0$, then we have:
%
\begin{equation*}
%
\frac{1}{N}\sum_{i=0}^{N-1} \varepsilon^2[i] \xrightarrow{a.s.} \sigma^2.
%
\end{equation*}

We now analyze the third-order autocorrelation. Let us fix $\ell_1\geq\ell_2\ge0$. We have:
%
\begin{align*}
%
&a_y^3[\ell_1,\ell_2] 
= \frac{1}{N}\sum_{i=0}^{N-1-\ell_1} y[i]y[i+\ell_1]y[i+\ell_2]
\nonumber \\
%
=& \underbrace{ \frac{ML}{N}\frac{1}{M}\sum_{j=1}^M 
	\frac{1}{L}\sum_{i=0}^{L-1-\ell_1}x[i]x[i+\ell_1]x[i+\ell_2]   }_{(1)}
\\& + \underbrace{\frac{1}{N}\sum_{i=0}^{N-1-\ell_1} \ep[i]\ep[i+\ell_1]\ep[i+\ell_2]}_{(2)}
\nonumber \\
&+ \underbrace{\frac{1}{N}\sum_{j=1}^{M} 
	\sum_{i=0}^{L-1} x[i]\ep[s_j + i+\ell_1]\ep[s_j+ i+\ell_2]}_{(3)}
\\& + \underbrace{\frac{1}{N}\sum_{j=1}^{M} 
	\sum_{i=0}^{L-1} \ep[s_j+i-\ell_1]x[i]\ep[s_j+ i+\ell_2-\ell_1]}_{(4)}
\nonumber \\
&+ \underbrace{\frac{1}{N}\sum_{j=1}^{M} 
	\sum_{i=0}^{L-1} \ep[s_j+i-\ell_2]\ep[s_j+i+\ell_1-\ell_2]x[i]}_{(5)}
\\ & + \underbrace{\frac{1}{N}\sum_{j=1}^{M} 
	\sum_{i=0}^{L-\ell_1+\ell_2-1} \ep[s_j+i-\ell_2]x[i+\ell_1-\ell_2]x[i]}_{(6)}
\nonumber \\
&+ \underbrace{\frac{1}{N}\sum_{j=1}^{M} 
	\sum_{i=0}^{L-\ell_2-1} x[i]\ep[s_j + i+\ell_1]x[i+\ell_2]}_{(7)}
\\ & + \underbrace{\frac{1}{N}\sum_{j=1}^{M} 
	\sum_{i=0}^{L-\ell_1-1} x[i]x[i+\ell_1]\ep[s_j+ i+\ell_2]}_{(8)}.
%
\end{align*}
%
Terms (6), (7) and (8) are linear in $\ep$, and can easily be shown to converge to 0 almost surely by the law of large numbers, by similar arguments as used previously. Term (1) converges to $\gamma a_x^3[\ell_1,\ell_2]$ almost surely, for the same reasons as~\eqref{eq:2nd_moment_signal_term}. To deal with terms (2)--(5), we must distinguish between different values of $\ell_1$ and $\ell_2$.

@@

{\bf Case 2:} $0=\ell_2 < \ell_1$. Term (2) is an average of products of the form $\ep[i]^2\ep[i+\ell_1]$, which have mean zero; consequently, term (2) converges to 0 almost surely. The same argument as for Case 1 shows that (3) and (5) also converge to 0. For term (4), we write:
%
\begin{align*}
%
&\frac{1}{N}\sum_{j=1}^{M} 
\sum_{i=0}^{L-1} \ep[s_j+i-\ell_1]x[i]\ep[s_j+ i+\ell_2-\ell_1]
\nonumber \\
&= \frac{ML}{N}\frac{1}{L}\sum_{i=0}^{L-1}x[i] \frac{1}{M}\sum_{j=1}^{M} \ep[s_j+i-\ell_1]^2
\\& \xrightarrow{a.s.} \gamma \frac{1}{L} \sum_{i=0}^{L-1}x[i] \sigma^2 = \gamma a_x^1 \sigma^2. \nonumber
%
\end{align*}

@@

Now let us compute the probability of failure conditional on the event $\eta = 0$. In this case, failure occurs when $X \in S$. Since $X|(\eta=0) \sim N(\theta_0,\sigma^2)$, we can write $X | (\eta=0) = \sigma Z + \theta_0$, where $Z \sim N(0,I)$. On that condition,

@@

\section{Proof of Proposition~\ref{prop:gamma_sigma}} \label{sec:proof_prop_gamma_sigma}
We will prove that both $\sigma$ and $\gamma$ are identifiable from the observed first three moments of $y$. For convenience, we will work with $\beta = \gamma / L$ rather than $\gamma$ itself. We will construct two quadratic equations satisfied by $\beta$ from observed quantities, independent of $\sigma$. Then, we will show that these equations are independent, and hence that $\beta$ is uniquely defined.  Given $\beta$, we can estimate $\sigma$ using Proposition~\ref{prop:gamma}.

@@

Recall that $a_y^1 = \beta(\one^Tx)$ and  
and $a_y^2[0] = \beta\|x\|^2+\sigma^2$, where $\one\in\RL$ stands for vector of ones. Taking the product:
\begin{equation}\label{eq:E1}
\begin{split}
E_1 &:= a_y^1a_y^2[0] =  (\beta(\one^Tx))(\beta\|x\|^2+\sigma^2) \\
& = \sigma^2a_y^1 + L\beta^2F_1,
\end{split}
\end{equation}
where $F_1 := a_x^3[0,0] + \sum_{j=1}^{L-1}(a_x^3[j,j] + a_x^3[0,j])$. 
The terms of $F_1$ can be also estimated from $a_y^3$, while taking the scaling and bias terms into account:
\begin{equation} \label{eq:E2}
E_2:= L\beta F_1 + (2L+1)\sigma^2a_y^1.
\end{equation}
Therefore, from~\eqref{eq:E1} and~\eqref{eq:E2} we get
\begin{equation} \label{eq:E12}
E_2\beta -(2L+1)\sigma^2\beta a_y^1 = E_1-\sigma^2a_y^1.
\end{equation}
Let $E_3:=a_y^2[0] + 2\sum_{j = 1}^{L-1}a_y^2[j]$ and recall from Proposition~\ref{prop:gamma}:
\begin{equation} \label{eq:sigma2}
\sigma^2 = E_3 - (a^1_y)^2/\beta. 
\end{equation} 
Plugging into~\eqref{eq:E12} and rearranging we get 
\begin{equation} \label{eq:quad1}
\mathcal{A}\beta^2 + \mathcal{B}\beta + \mathcal{C} = 0,
\end{equation}
where 
\begin{align*}
\mathcal{A} &= E_2 - (2L+1)a_y^1E_3, \\ 
\mathcal{B} &= -E_1 + (2L+1)(a_y^1)^3 + a_y^1E_3  , \\
\mathcal{C} &= -(a_y^1)^3.
\end{align*}
Importantly, these coefficients are observable quantities.

Next, we derive the second quadratic equation for $\beta$. We notice that 
\begin{equation} \label{eq:E3}
E_4  = \frac{1}{L}(a_y^1)^3 = \frac{1}{L}\beta^3 (\one ^Tx)^3   = \beta^3 F_2,
\end{equation}
where 
\begin{equation*}
F_2 =  a_x^3[0,0] + 3\sum_{j=1}^{L-1}a_x^3[j,j] + 3\sum_{j=1}^{L-1}a_x^3[0,j] + 6\sum_{1\leq i < j\leq L-1}a_x^3[i,j].
\end{equation*}
On the other hand, from $a_y^3$ we can directly estimate $F_2$ up to scale and bias
\begin{equation*} \label{eq:E4}
E_5 = L \beta F_2 + (6L-3)\sigma^2a_y^1.
\end{equation*}
Taking the ratio:
\begin{equation*} 
\frac{E_5}{E_4} = \frac{L}{\beta^2} + \frac{(6L-3)\sigma^2a_y^1}{E_4}, 
\end{equation*}
we conclude:
\begin{equation*}
\sigma^2 = \frac{E_5}{a_y^1(6L-3)}  - \frac{LE_4}{\beta^2a_y^1(6L-3)}.
\end{equation*}
Using~\eqref{eq:sigma2} and rearranging we get the second quadratic:
\begin{equation} \label{eq:quad2}
\mathcal{D}\beta^2 + \mathcal{E}\beta + \mathcal{F} = 0,
\end{equation}
where
\begin{align*}
\mathcal{D} &= E_3 - \frac{E_5}{a_y^1(6L-3)}, \\ 
\mathcal{E} &= -(a_y^1)^2, \\
\mathcal{F} &= \frac{LE_4}{a_y^1(6L-3)}.
\end{align*}

To complete the proof, we need to show that the two quadratic equations~\eqref{eq:quad1} and~\eqref{eq:quad2} are independent. To this end, it is enough to show that the ratio between the coefficients is not the same. 
From~\eqref{eq:quad1} and~\eqref{eq:E1}, we have 
\begin{equation*}
\begin{split}
\frac{\mathcal{B}}{\mathcal{C}} &= \frac{E_1 - (2L+1)(a_y^1)^3 - a_y^1E_3}{(a_y^1)^3} \\&= \frac{a_y^2[0] - (2L+1)(a_y^1)^2 - E_3}{(a_y^1)^2}.
\end{split}
\end{equation*}
In addition, using~\eqref{eq:E3}
\begin{equation*}
\frac{\mathcal{E}}{\mathcal{F}} = \frac{(3-6L)(a_y^1)^3}{LE_4} = 3 - 6L . 
\end{equation*}

Now, suppose that the quadratics are dependent. Then, $\frac{\mathcal{B}}{\mathcal{C}} =\frac{\mathcal{E}}{\mathcal{F}} $, or, 	
\begin{equation*}
a_y^2[0] - (2L+1)(a_y^1)^2 - E_3 = (a_y^1)^2(3-6L).
\end{equation*}
Rearranging the equation and writing in terms of $x$ we get 
\begin{equation} \label{eq:cond}
4(L-1)\beta (\1^\top x)^2  - (\1^\top x)^2 + \|x\|^2 = 0.
\end{equation}	
For generic $x$,  this polynomial equation is not satisfied and the equations are independent. 
Furthermore, from the inequality $L\|x\|^2 \ge (\1^\top x)^2$ it follows immediately that the equations must be independent so long as
\begin{equation*}
\beta \geq \frac{1}{4L}.
\end{equation*}
